---
title: "Current Account and Balance of Trade"
author: "Sebastian Krantz"
date: "6/24/2021"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)

```

```{css}
/*
body{
  background-color: #959a9c;
}

h2{
  font-weight: bold;
}
*/
```

```{r, data, include=FALSE}
library(ugatsdb)
ugatsdb_reconnect()
bop_data <- get_data("BOU_BOP", c("CAB", "CA_G"))
names(bop_data)[-1] <- c("Current Account", "Balance of Trade")
```

```{r constants, include=FALSE}
mean_CAB <- mean(bop_data$`Current Account`)
mean_TB <- mean(bop_data$`Balance of Trade`)

text <- if(mean_CAB > mean_TB) "higher" else "lower"
```

## A Static Plot

The plot below shows the quarterly **current account balance** and **balance of trade for Uganda**.
The average CAB over the period has been **`r round(mean_CAB)`** whereas the average trade balance has been **`r round(mean_TB)`**. Thus on average the CAB has been `r if(mean_CAB > mean_TB) "higher" else "lower"` than the TB. 

```{r static}
library(xts)
bop_xts <- as.xts(bop_data)
plot(bop_xts, legend.loc = "topright", 
     main = "Current account and balance of trade",
     yaxis.right = FALSE)

```

You can see that both move together very closely. We can observe:

* CAB is initially a lot higher than TB
* CAB deteriorates and is about equal to TB in 2010
  + Yet the CAB is more volatile than the TB

## A Dynamic Plot

This shows a dynamic plot of the same thing. 

```{r dynamic}
library(dygraphs)
dygraph(bop_xts)
```
 
### A table 

```{r, results='asis'}

knitr::kable(tail(bop_data))
```


