---
title: "COVID-Uganda"
author: "Sebastian Krantz"
date: "6/24/2021"
output: beamer_presentation
    
header-includes:
  \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r data, include=FALSE}
ugatsdb::ugatsdb_reconnect()
data <- ugatsdb::get_data("MOF_POE", c("CIEA", "BTI", "PMI", "EX", "IM", "TB"), from = 2019)
data_xts <- xts::as.xts(data)
library(collapse)
lab <- vlabels(data)
```

## Indicators

The impact of the COVID pandemic on Uganda can be seen through the following indicators.

```{r table}
library(kableExtra)
library(magrittr)
qsu(data, cols = -1, vlabels = TRUE) %>% 
  round(2) %>%
   kbl(format = "latex", booktabs = TRUE, linesep = "", 
       caption = "Key Economic Indicators") %>% 
     kable_styling(latex_options = "scale_down")
```

**PMI** and **BTI** are indices with a threshold at 50, below which economic 
activity is considered to be low. The **CIEA** is an index composed of increasing variables. 

## Composite Index of Economic Activity

- The **CIEA** shows a sharp drop between February and July 2020, when Uganda had the first lockdown. 

```{r}
plot(data_xts[, "CIEA"], col = "orange", main = lab["CIEA"])
```

## Business Tendency Indicator

- The **BTI** shows a sharp drop in Fbruary 2020, and has not quite recovered yet in spring 2021.  

```{r}
plot(data_xts[, "BTI"], col = "darkgreen", main = lab["BTI"])
```

## Purchasing Managers Index

- The **PMI** shows a sharp drop in between February and September 2020.  

```{r}
plot(data_xts[, "PMI"], col = "magenta", main = lab["PMI"])
```

## Exports and Imports

- Exports and Imports declined between May and August 2020.   

```{r}
plot(data_xts[, c("EX", "IM")], main = "Exports and Imports (US$ millions)", legend.loc = "topleft")
```

## Balance of Trade 

- The trade balance has improved slightly suring the lockdown, as other countries did not have such a stringent lockdown.    

```{r}
plot(data_xts[, "TB"], col = "red", main = lab["TB"])
```

## Conclusion

- COVID-19 and the first lockdown between March 2020 and July 2020 had a profound impact on the Ugandan economy. 
